NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-06-25 Manuscript/manuscriptPS.tex
Update text

Previous experimental studies have concluded that
PS headgroups have more rigid conformation than
phophocholines (PC) due to hydrogen bonding or
electrostatic interactions \cite{browning80,buldt81}.
Multivalent cations and Li$^+$ are able to form strong
dehydrated molecular complexes with PS lipids,
while monovalent ions interact more weakly with PS
containing bilayers \cite{hauser77,kurland79,hauser83,hauser85,feigenson86,mattai89,roux90,roux91}.
Dilution of bilayers with PC lipids makes PS headgroups
less rigid and reduces propensity for the formation of
strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.

The structure of PS lipid headgroups and their 
interactions with ions have been studied with
various experimental methods and theoretical techniques \cite{roux90,melcrova16,??}.
However, the consesus has not been reached 
due to the difficulties to interpret the experimental data \cite{??} and
the inaccuracies in simulation models at the
headgroup region \cite{botan15,catte16,ollila16}.
Some studies propose that the negatively charged lipids
attract cations only due to the increase of local concentration
in the vicinity of membranes and that the binding constant of cations is similar
to zwitterionic and negatively charged lipids \cite{seelig90,sinn06,??}.
On the other hand, some studies propose specific binding of calcium directly to PS lipid
headgroups \cite{vernier09,boettcher11,??},
The NMR data proposes that the PS headgroup is more rigid than PC,PE or PG
headgroups, but more detailed interperetation has not been done.

>>>>

Previous experimental studies have concluded that
PS headgroups are more rigid than phophocholines (PC)
due to the hydrogen bonding network or
electrostatic interactions \cite{browning80,buldt81}.
Multivalent cations and Li$^+$ are able to form strong
dehydrated molecular complexes with PS lipids,
while monovalent ions interact more weakly with PS
containing bilayers \cite{hauser77,kurland79,hauser83,hauser85,feigenson86,mattai89,roux90,roux91,boettcher11}.
On the other hand, some studies propose that the specific binding
affinity is similar to the negatively charged and zwitterionic lipids and that
the increased cation binding to negatively charged lipid bilayer arise only due
to the increase of local cation concentration in the vicinity of membranes \cite{seelig90,sinn06}.
Dilution of bilayers with PC lipids makes PS headgroups
less rigid and reduces propensity for the formation of
strong complexes with multivalent ions \cite{browning80,buldt81,roux90,roux91}.
The molecular level interpretation of these observations is,
however, not available.

Several classical molecular dynamics (MD) simulation studies are done
to understand PS headgroups, their influence on lipid bilayer properties and
interactions with ions \cite{vernier09,boettcher11,melcrova16,??}.
However, the consesus has not been reached 
due to the difficulties to interpret the experimental data \cite{??} and
the inaccuracies in simulation models at the
headgroup region \cite{botan15,catte16,ollila16}.
